Bipartite Fidelity and Loschmidt Echo of Bosonic Conformal Interface

### 2017-05-11 F_correction.tex
[draft] appendix done; analytic results done; discussion draft 1.

\todo[inline]{cite main text figure for conformal transformation}
In the course of derivation the free energy subject to various boundary conditions, we use conformal transformation to convert the space time diagram with slits to a cylinder diagram, where the boundary state (and ground state energy calculation in App.~\ref{app:gnd_dn_lambda}) is applicable. However, free energy, especially its constant part, is not invariant under the conformal transformation, because the boundaries partially break the conformal symmetry. In this Appendix, we point out two corrections -- one from the outer boundary and the other from the inhomogeneous Schwartzian term -- to get the correct exponent of the fidelity and Loschmidt echo.

% outer boundary correction
It is discussed in the pioneering work\cite{cardy_finite-size_1988} by Cardy and Peschel that boundary will contribute logarithmic term, which manifest itself as the geodesic curvature integral in the trace anomaly. Here we demonstrate this principle using the disk free energy example in Ref.~\cite{cardy_finite-size_1988}. Consider an annulus on flat space with inner radius $r_1$ and outer radius $r_2$. Its free energy is
\begin{equation}
F({\rm annulus}) = -  \frac{c}{6} \ln \frac{r_2}{r_1} 
\end{equation}
On the other hand the free energy of a disk of radius $r_2$ is
\begin{equation}
  F( {\rm disk} ) = - \frac{c}{6} \ln \frac{r_2}{a}
\end{equation}
where $a$ is the short distance regulator. The disk free energy is completely contributed by its outer boundary and so the one can interpret the annulus free energy as the additive contribution from its outer and inner surfaces (which is also consistent with the geodesic curvature integral calculation),
\begin{equation}
F( {\rm annulus} ) = -  \frac{c}{6} \ln \frac{r_2}{a} +  \frac{c}{6} \ln \frac{r_1}{a} = - \frac{c}{6} \ln \frac{r_2}{r_1}
\end{equation}
An annulus becomes a disk when its inner radius is of order $a$, and we can see that the contribution from the inner surface $\frac{c}{6} \ln \frac{r_1}{a}$ becomes negligible.

\todo[inline]{citing figure}
We now scrutinize the conformation transformation from the slit diagram to the annulus. In the slit diagram, the regulators all have radius that is order of the short distance cut-off. They will have negligible contributions to the free energy. However, after doing the conformal transformation (passive), the outer surface will contribute $-\frac{c}{6} \ln \frac{r_2}{r_1}$ which should not be there. Using the parameters in the free energy calculation in other appendices, we have 
\begin{equation}
F_{\xi } = F_{z} + \frac{c}{6} \beta 
\end{equation}

% staircase geometry correction
The annulus is called staircase geometry Ref.~\onlinecite{cardy_finite-size_1988} due to its angular direction of time evolution. The traditional radial quantization however has radial direction to be the time direction. One can go through a series of conformal maps, one can show that that the Hamiltonian of the staircase and rectangle has a shift due to the Schwartzian
\begin{equation}
H_{z} = H_{\xi} - \frac{c}{24\pi} \beta 
\end{equation}
After the evolution for $2\pi$, the difference in the free energy is
\begin{equation}
F_w = F_{\xi} - \frac{c}{12} \beta 
\end{equation}
Gathering the two together, we obtain the missing $\frac{c}{12} \beta$ between the echo diagram and the Casimir energy, 
\begin{equation}
F_{w} = F_z + {\color{red} \frac{c}{12}\beta }
\end{equation}
\begin{figure}[h]
\centering
\includegraphics[width=\textwidth]{fig_uncon_rq.pdf}
\caption{Unconventional radial quantization on annulus(torus). Left panel: unlike the conventional radial quantization, we view the dashed line as a state and to time evolution in the radial direction. The blue circles are regarded as boundaries conditions rather than asymptotic states. Middle panel: $\xi = \ln w$ plane. Right panel: $z = \exp(i \frac{2\pi}{\beta} \xi)$ plane. We treat those dashed lines as asymptotic states and perform standard radial quantization here.}
\label{fig:uncon_rq}
\end{figure}

\todo[inline]{There are two series of conformal transformation, one is the main text, the other is the staircase geometry. Both use $w, z, \xi$, I will fix the notation after the main text figure is complete. Maybe detailed computation of the Schwartzian term is necessary, because of the sign. }

>>>>

In the course of derivation the free energy subject to various boundary conditions, we use conformal transformation to convert the spacetime diagram with slits to a cylinder diagram, where the boundary state (and ground state energy calculation in App.~\ref{app:gnd_dn_lambda}) is applicable. However, free energy is not invariant under the conformal transformation, because the boundaries partially break the conformal symmetry. In this Appendix, we point out two corrections -- one from the outer boundary regulator and the other from the inhomogeneous Schwartzian term -- to get the correct exponent of the fidelity and Loschmidt echo.

% outer boundary correction
It is discussed in Cardy and Peschel's pioneering work\cite{cardy_finite-size_1988} that boundary will contribute a logarithmic term in the free energy. The original of these term comes from the nonzero geodesic curvature of the boundary, which when adding back the bulk curvature term constitutes the trace anomaly. Here we demonstrate this principle using the disk free energy example in Ref.~\onlinecite{cardy_finite-size_1988}. Consider an annulus on flat space with inner radius $r_1$ and outer radius $r_2$. Its free energy is
\begin{equation}
F({\rm annulus}) = -  \frac{c}{6} \ln \frac{r_2}{r_1} 
\end{equation}
On the other hand the free energy of a disk of radius $r_2$ is
\begin{equation}
  F( {\rm disk} ) = - \frac{c}{6} \ln \frac{r_2}{a}
\end{equation}
where $a$ is the short distance regulator. The disk free energy is completely contributed by its outer boundary with other parts being conformal invariant. One can then interpret the annulus free energy as an additive contribution from its outer and inner surfaces as (which is also consistent with the geodesic curvature integral calculation),
\begin{equation}
F( {\rm annulus} ) = -  \frac{c}{6} \ln \frac{r_2}{a} +  \frac{c}{6} \ln \frac{r_1}{a} = - \frac{c}{6} \ln \frac{r_2}{r_1}
\end{equation}
An annulus becomes a disk when its inner radius is of order $a$, and we can see that the contribution from the inner surface, because $\frac{c}{6} \ln \frac{r_1}{a}$ becomes negligible compared to the other term.

A similar outer surface logarithmic term also appears in the conformation transformation from the slit diagram to the annulus in Fig.~\ref{fig:H-tau_fold}. In the slit diagram, the regulators all have radii that are at the order of the short distance cut-off. They will have negligible contributions to the free energy. However, after doing the conformal transformation, the outer surface will contribute $-\frac{c}{6} \ln \frac{r_2}{r_1}$ which should not be there. Using the cylinder parameters in App.~\ref{app:lambda_12}, the $\xi$ plane and $z$ plane free energy are related through
\begin{equation}
F_{\xi } = F_{z} + \frac{c}{6} \beta 
\end{equation}

% staircase geometry correction
The annulus on the $\xi$ plane is called staircase geometry in Ref.~\onlinecite{cardy_finite-size_1988} due to its angular time evolution. The traditional radial quantization however has radial direction to be the time. One can show that the Hamiltonian of the staircase and rectangle has a shift due to the Schwartzian\cite{cardy_finite-size_1988}
\begin{equation}
H_{w} = H_{\xi} - \frac{c}{24\pi} \beta 
\end{equation}
After the evolution for $2\pi$ (in the folding picture, the evolution is only $\pi$ but there are two bosons), the difference in the free energy is
\begin{equation}
F_w = F_{\xi} - \frac{c}{12} \beta 
\end{equation}
Gathering the two terms, we obtain the missing correction $\frac{c}{12} \beta$ between the slit and cylinder diagram, 
\begin{equation}
F_{w} = F_z + \frac{c}{12}\beta
\end{equation}

### 2017-05-11 analy_eval.tex
[draft] appendix done; analytic results done; discussion draft 1.

For the process ($c$ is assumed to be the same as $a$)
\begin{equation}
S_i( \theta_1 ) \rightarrow S_j( \theta_2 ) 
\end{equation}
the free energy is
\begin{equation}
\mathcal{F}( \beta )  = 
\left\lbrace
\begin{aligned}
  &\frac{1}{2}(-|x| + x^2 )\beta  \quad &i = j \\
  &\frac{1}{16}\beta   \quad &i \ne j   \\
\end{aligned} \right. \quad x = \frac{\theta_2 - \theta_1}{\pi} 
\end{equation}
We can then read off the fidelity and Loschmidt echo power law decay exponent with the identification $\beta = 2 \ln L$ and $ 4 \ln \tau$.

>>>>

For the process ($c$ is assumed to be the same as $a$)
\begin{equation}
S_i( \theta_1 ) \rightarrow S_j( \theta_2 ) 
\end{equation}
the free energy is
\begin{equation}
\mathcal{F}( \beta )  = 
\left\lbrace
\begin{aligned}
  &\frac{1}{2}(|x| - x^2 )\beta  \quad &i = j \\
  &\frac{1}{16}\beta   \quad &i \ne j   \\
\end{aligned} \right. \quad x = \frac{\theta_2 - \theta_1}{\pi} 
\end{equation}
We can then read off the fidelity and Loschmidt echo power law decay exponent with the identification $\beta = 2 \ln L$ and $ 4 \ln \tau$.

As analyzed in Sec.~\ref{sec:analytic_numerics}, $S_2( \theta)$ interpolates between DD and NN, $S_1( \theta )$ interpolates between DN and ND. In the region accessible to the numerical calculation in the lattice model, we choose the process 
\begin{equation}
{\rm DD} + {\rm DD} \rightarrow  \lambda + {\rm DD}
\end{equation}
to verify
\begin{equation}
\mathcal{F} = 
\left\lbrace
\begin{aligned}
\frac{1}{8}\ln L  &\quad\text{fidelity}  \\
\frac{1}{4}\ln t   &\quad \text{echo}   \\
\end{aligned} \right.  
\end{equation}
where same results have already been obtained for $\lambda = {\rm P}$\cite{stephan_logarithmic_2013,stephan_local_2011,vasseur_universal_2014,vasseur_crossover_2013,kennes_universal_2014}. Another process
\begin{equation}
{\rm DN} + {\rm DN} \rightarrow \lambda + {\rm DN} 
\end{equation}
is used to verify
\begin{equation}
\mathcal{F} = 
\left\lbrace
\begin{aligned}
 (x - x^2 )\ln L   &  \quad {\rm fidelity} \\
 2(x - x^2 ) \ln t  & \quad \text{ Loschmidt echo} \\
\end{aligned} \right. 
\end{equation}
where $\lambda = \tan \theta$ and $x = \frac{\theta}{\pi}$.

### 2017-05-11 bd_notation.tex
[draft] appendix done; analytic results done; discussion draft 1.

We use chemical reaction equation style to represent the change of boundary conditions. Taking the example of the echo diagrams in Fig.~{\color{red} TODO}, there are three boundary conditions $a,b,c$ in the folding picture, which represents status on two ends of the chain before and after the quench. The choice of a uniform $c$ boundary condition on the far end of the chain is to isolate the effect coming from the bcc on the conformal interface. The process $a + c \rightarrow b + c$ represents the change of boundary condition form the combination $a$/$c$ on the two ends to $b$/ $c$ after the quench. Since each letter can take a general conformal interface defined by the $S$ matrix in Eq.~\eqref{eq:S1_S2}, we denote it as
\begin{equation}
S_a( \theta_a ) + S_c( \theta_c) \rightarrow S_b( \theta_b )  + S_c( \theta_c ) 
\end{equation}
Analytic computation can be performed when we neglect the boundary condition $c$, assuming $L \gg \tau$. To remove the bcc operator from $c$ to $a$ at infinity, we usually take $a = c$ in the numerical calculation.

In the "cut-and-join" protocol we considered, $a$ should be one of 'DD', 'DN', 'ND', 'NN', $b$ is taken to be $S_1( \theta )$ or $S_2( \theta )$. The physical situation of connecting two compact bosons and our lattice model corresponds to the choice of $S_1( \theta)$, we also reserve the notation $\lambda$ for this type of the boundary condition. For instance, the notation for the process presented in Fig.~{\color{red} TODO} is
\begin{equation}
{\rm DN }  + {\rm DN} \rightarrow \lambda + { \rm DN} 
\end{equation}
The scaling dimension of this bcc operator does indeed deviate from the completely transmitting case for certain choice of boundary conditions.

>>>>

We use chemical reaction style to represent the change of boundary conditions. Taking the example of the echo diagrams in Fig.~\ref{fig:echo}, there are three boundary conditions $a,b,c$ in the folding picture, which represents status on two ends of the chain before and after the quench. The choice of a uniform $c$ boundary condition on the far end of the chain is to isolate the effect coming from the bcc on the conformal interface. The process $a + c \rightarrow b + c$ represents the change of boundary condition form the combination $a$/$c$ on the two ends to $b$/ $c$ after the quench. Since each letter can take a general conformal interface defined by the $S$ matrix in Eq.~\eqref{eq:S1_S2}, we denote it as
\begin{equation}
S_a( \theta_a ) + S_c( \theta_c) \rightarrow S_b( \theta_b )  + S_c( \theta_c ) 
\end{equation}
Analytic computation can be performed when we neglect the boundary condition $c$, assuming $L \gg \tau$. To remove the bcc operator from $c$ to $a$ at infinity, we usually take $a = c$ in the numerical calculation.

In the "cut-and-join" protocol we considered, $a$ should be one of 'DD', 'DN', 'ND', 'NN', $b$ is taken to be $S_1( \theta )$ or $S_2( \theta )$. The physical situation of connecting two compact bosons and our lattice model corresponds to the choice of $S_1( \theta)$, we also reserve the notation $\lambda$ for this type of the boundary condition. For instance, the notation for the process presented in Fig.~\ref{fig:echo} is
\begin{equation}
{\rm DN }  + {\rm DN} \rightarrow \lambda + { \rm DN} 
\end{equation}
The scaling dimension of this bcc operator does indeed deviate from the completely transmitting case for certain choice of boundary conditions.
